Modelling BCG vaccination in the UK

### 2019-07-04 rmds/02-background.Rmd
hannah comments

Tuberculosis (TB) is thought to infect over 1.7 billion people globally, of which 5-15% will develop active TB in their lifetime.[@TheWorldHealthOrganization:2018va] Of this number around 10% are likely to die from TB or TB related causes. TB is preventable and curable, but the majority of cases occur in less economically developed countries and are never diagnosed.[@TheWorldHealthOrganization:2018va] In the following section, the natural history of TB, the risk factors, the treatment, the global impact, and the impact in England and Wales are explored.

>>>>

Tuberculosis (TB) is thought to infect over 1.7 billion people globally, of which 5-15% will develop active TB in their lifetime.[@TheWorldHealthOrganization:2018va] Of this number around 10% are likely to die from TB or TB related causes.[@TheWorldHealthOrganization:2018va] TB is preventable and curable, but the majority of cases occur in less economically developed countries and are never diagnosed.[@TheWorldHealthOrganization:2018va] In the following section, the natural history of TB, the risk factors, the treatment, the global impact, and the impact in England and Wales are explored.

### 2019-07-04 rmds/index.Rmd
hannah comments

The BCG vaccine was developed in 1921 and was introduced to the UK in 1953. Globally, it has been shown to offer variable protection that may reduce over time.[@TheWorldHealthOrganization:2018va] However, there is strong evidence that BCG offers high levels of protection for children, and more generally within the UK born population.[@Roy2014] It remains the only licensed TB vaccine with over 100 million doses given globally each year. Serious side effects are rare but scarring commonly occurs at the site of injection. In 2005, the UK withdrew the universal BCG program for those at school age and introduced a targeted program of vaccination for babies that were deemed to be at high risk.[@Zwerling2011a] This was motivated by several years of declining transmission, the evidence of high levels of protection in children and a belief that other control measures would be more cost-effective.[@Zwerling2011a] Since this change in policy, declining incidence appears to support this decision.[@PHE2017] However, due to TB’s complex dynamics, the long-term effects are difficult to predict.

@@

In this thesis, a variety of statistical models are used to explore complex multivariate relationships. Uses cases include: adjusting for confounding variables when estimating the relationship between BCG vaccination and TB outcomes (Chapter \@ref(beneficial-bcg-out)); and estimating the impact on incidence rates from ending the BCG schools scheme after accounting for various confounders (Chapter \@ref(direct-eff)).

@@

  - To describe the current epidemiology of TB in England, in the context of global TB epidemiology.
  - To assess some of the statistical modelling evidence used to justify the 2005 change in BCG vaccination policy in the UK.
  - To assess whether there is evidence in routinely-collected surveillance data that BCG vaccination impacts outcomes for TB cases in England.
  - To assess the effects of the 2005 change in vaccination policy on those eligible for vaccination.
  - To develop a parsimonious transmission dynamic model of TB which captures current, and historic, vaccination policy and reflects our current understanding of TB epidemiology in England.
  - To fit this model using all available data sources.
  - To investigate the effectiveness of universal school-age vs. universal neonatal vs. no vaccination using the previously developed transmission dynamic model.

@@

* **Chapter \@ref(discussion-chapt):** Results from all previous Chapters are summarised and discussed as a whole. The strengths and weaknesses of the analysis in this thesis are outlined. Further work is suggested. Future work is outlined.

>>>>

The Bacillus Calmette–Guérin (BCG) vaccine was developed in 1921 and was introduced to the UK in 1953. Globally, it has been shown to offer variable protection that may reduce over time.[@TheWorldHealthOrganization:2018va] However, there is strong evidence that BCG offers high levels of protection for children, and more generally within the UK born population.[@Roy2014] It remains the only licensed TB vaccine with over 100 million doses given globally each year. Serious side effects are rare but scarring commonly occurs at the site of injection. In 2005, the UK withdrew the universal BCG program for those at school age and introduced a targeted program of vaccination for babies that were deemed to be at high risk.[@Zwerling2011a] This was motivated by several years of declining transmission, the evidence of high levels of protection in children and a belief that other control measures would be more cost-effective.[@Zwerling2011a] Since this change in policy, declining incidence appears to support this decision.[@PHE2017] However, due to TB’s complex dynamics, the long-term effects are difficult to predict.

@@

In this thesis, a variety of statistical models are used to explore complex multivariate relationships. Uses cases include: adjusting for confounding variables when estimating the relationship between BCG vaccination and TB outcomes (Chapter \@ref(benefical-bcg-out)); and estimating the impact on incidence rates from ending the BCG schools scheme after accounting for various confounders (Chapter \@ref(direct-eff)).

@@

  - To describe the current epidemiology of TB in England, in the context of global TB epidemiology.
  - To assess some of the statistical modelling evidence used to justify the 2005 change in BCG vaccination policy in the UK.
  - To assess whether there is evidence in routinely-collected surveillance data that BCG vaccination impacts outcomes for TB cases in England.
  - To assess the effects of the 2005 change in vaccination policy on those eligible for vaccination.
  - To develop a parsimonious transmission dynamic model of TB that captures current, and historic, vaccination policy and reflects our current understanding of TB epidemiology in England.
  - To fit this model using all available data sources.
  - To investigate the effectiveness of universal school-age vs. universal neonatal vs. no vaccination using the previously developed transmission dynamic model.

@@

* **Chapter \@ref(discussion-chapt):** Results from all previous Chapters are summarised and discussed as a whole. The strengths and weaknesses of the analysis in this thesis are outlined. Further work is outlined.
